Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-05-27 publication_materials/ms_rbc.tex
timeout

This approximation is valid in equilibrium, but in general, we would obtain undesired $z$-dependence in $A^2$.
Though we might be tempted to take a norm, the best way to account for this $z$-dependence, is to make an analogous argument with respect to the growth rates
\begin{align}
    \sigma_{\rm{diff}} - \sigma &\approx -A^2(\sigma_{\rm{adv}} - \sigma).
\end{align}
where $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ are the growth rates computed from $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ respectively.
$\sigma$ vanishes due to marginal stability.
Later we will see that this approach can be easily generalized to account for multiple simultaneously marginal modes.

In practice, we initialize $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ with $\overline{T}(z, t_0)$, and evolve the equations to $t=t_0+\Delta t$.
Suppose $k_0$ is the wavenumber of the marginally-stable mode.
Then we compute $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ at $k_x = k_0$ and $t=t_0+\Delta t$.
A preliminary approximation $\Lambda^2_0$ of the appropriate amplitude $A^2$ is given by
\begin{equation}
    A^2 \approx \Lambda^2_0 \equiv -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where the square in $A^2$ and $\Lambda^2_0$ is due to the quadratic nonlinearity of the eigenfunctions in the advective term and the minus sign is due to the fact that $\sigma_{\rm{adv}} < 0$ and $\sigma_{\rm{diff}} > 0$ in general.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
The subscript 0 denotes a preliminary estimate which we later refine.
In general, when $\Lambda^2_0 > 0$, the root-finding algorithm is stable.
The alternative can occur when the timestep approaches machine precision, in which case $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ are subject to floating point errors of comparable order to themselves.

>>>>

This approximation is valid in equilibrium, but in general, we would obtain undesired $z$-dependence in $A^2$.
Though we might be tempted to take a norm, the best way to account for this $z$-dependence, is to make an analogous argument with the growth rates
\begin{align}
    \sigma_{\rm{diff}} - \sigma &\approx -A^2(\sigma_{\rm{adv}} - \sigma).
\end{align}
where $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ are computed from $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ respectively.
$\sigma$ vanishes due to marginal stability.
Later we will see that this approach can be easily generalized to account for multiple simultaneously marginal modes.

In practice, we initialize $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ at $t = t_0$, and evolve the equations to $t=t_0+\Delta t$.
Suppose $k_0$ is the wavenumber of the marginally-stable mode.
Then we compute $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ at $k_x = k_0$ and $t=t_0+\Delta t$.
A preliminary approximation $\Lambda^2_0$ of the appropriate amplitude $A^2$ is given by
\begin{equation}
    A^2 \approx \Lambda^2_0 \equiv -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where the square in $A^2$ and $\Lambda^2_0$ is due to the quadratic nonlinearity of the eigenfunctions in the advective term and the minus sign is due to the fact that $\sigma_{\rm{adv}} < 0$ and $\sigma_{\rm{diff}} > 0$ in general.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
The subscript 0 denotes a preliminary estimate which we later refine.
In general, when $\Lambda^2_0 > 0$, the root-finding algorithm is stable.
The alternative can occur when the timestep approaches machine precision, in which case $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ are subject to floating point errors of comparable order to themselves.
